Competing Bandits:\\ The Perils of Exploration under Competition \thanks{This is a merged and final version of two conference papers, \citet{CompetingBandits-itcs18} and \citet{CompetingBandits-ec19}, with a unified and streamlined presentation.}

### 2020-07-13 journal paper/content/sec-theory.tex
Added "monotonicity of alg2" proof.

\xhdr{Notation.} Let $\rew_i(n)$ denote the Bayesian-expected reward of principal $i$ at local step $n$. For a global round $t$, let $n_i(t)$ denote the number of global rounds before $t$ in which principal $i$ is chosen. Note that
    \[ \PMR_i(t):=  \E\sbr{ r_t\mid i_t = i }
        = \E\sbr{ \rew_i(n_i(t)+1) }. \]

@@

We characterize the inherent quality of an MAB algorithm in terms of its \emph{Bayesian Instantaneous Regret} (henceforth, \BIR), a standard notion from machine learning:
\begin{align}
\BIR(n) := \E_{\mu\sim\priorMu}
    \left[ \max_{a\in A} \mu_a\right] - \rew(n),
\end{align}
where $\rew(n)$ is the Bayesian-expected reward of the algorithm for the $n$-th step, when the algorithm is run in isolation. We are primarily interested in how \BIR scales with $n$: typically, how fast \BIR decreases with $n$. We treat $K$, the number of arms, as a constant unless specified otherwise. Let $\BIR_i(n) = \BIR(n)$ for the algorithm of principal $i$.

@@

%It is easy to see that \alg[2] is monotone. Indeed,

@@

As in the previous subsection, we can state a game-theoretic corollary for the \FiniteGame (as in Definition~\ref{def:restricted-competition}). We need a mild technical assumption that cumulative Bayesian regret (\BReg) tends to infinity. \BReg is a standard notion from the literature (along with \BIR):
\begin{align}\label{eq:SoftMax-BReg}
\BReg(n) := n\cdot \E_{\mu\sim\priorMu}
    \left[ \max_{a\in A} \mu_a\right] - \sum_{n=1}^n \rew(n')
    = \sum_{n'=1}^n \BIR(n').
\end{align}

>>>>

\xhdr{Notation.} Let $\rew_i(n)$ denote the realized reward of principal $i$ at local step $n$. For a global round $t$, let $n_i(t)$ denote the number of global rounds before $t$ in which principal $i$ is chosen. Note that
    \[ \PMR_i(t):=  \E\sbr{ r_t\mid i_t = i }
        = \E\sbr{ \rew_i(n_i(t)+1) }. \]

@@

We characterize the inherent quality of an MAB algorithm in terms of its \emph{Bayesian Instantaneous Regret} (henceforth, \BIR), a standard notion from machine learning:
\begin{align}\label{eq:theory-BIR-defn}
\BIR_i(n) := \E\sbr{ n\cdot \max_{a\in A} \mu_a - \rew_i(n)}.
\end{align}
We are primarily interested in how \BIR scales with $n$: typically, how fast \BIR decreases with $n$. We treat $K$, the number of arms, as a constant unless specified otherwise. 
%Let $\BIR_i(n) = \BIR(n)$ for the algorithm of principal $i$.

@@

Next, let us argue that \alg[2] is monotone. We spell out the proof for the sake of intuition. For each $n\in\N$, let $M_n$ be the number of times \alg[1] is invoked in the first $n$ steps of \alg[2], and let $\BIRgr_1(n)$ be the $\BIR$ of the ``greedy choice" after $n-1$ steps of \alg[1]. Let 
    $\mathtt{reg}_2(n) = n\cdot \max_a \mu_a - \rew_2(n)$ 
be the (frequentist) instantaneous regret of \alg[2]. Then by definition of the greedy modification it holds that 
\begin{align}\label{eq:random-BIR}
\E\sbr{ \mathtt{reg}_2(n) \mid M_n = m}
    &= (1-p) \cdot \BIR_1(m) + p \cdot \BIRgr_1(m) .\\
\BIR_2(n)
    &= \E\sbr{ (1-p) \cdot \BIR_1(M_n) + p \cdot \BIRgr_1(M_n) }.
\end{align}
Now, monotonicity of \alg[2] follows from monotonicity of $\BIR_1(\cdot)$ and $\BIRgr_1(\cdot)$.

@@

As in the previous subsection, we can state a game-theoretic corollary for the \FiniteGame (as in Definition~\ref{def:restricted-competition}). We need a mild technical assumption that cumulative Bayesian regret (\BReg) tends to infinity. \BReg is a standard notion from the literature (along with \BIR):
\begin{align}\label{eq:SoftMax-BReg}
\BReg(n) := n\cdot \E_{\mu\sim\priorMu}
    \left[ \max_{a\in A} \mu_a\right] - 
    \E\sbr{ \sum_{m=1}^n \rew(m)}
    = \sum_{m=1}^n \BIR(m).
\end{align}
